Application-Specific Enhanced Morphological Analyzer \\for Arabic Text Mining

### 2011-02-25 trunk/tex/sarf.aclhlt.tex
left one page and 4 lines to reduce...

\begin{document}
\maketitle
\begin{abstract}
Natural language processing (NLP) uses morphological 
analysis to abstract and annotate text.
Often times, 
%      and in particular in the context of the Arabic language, 
annotations resulting from 
a branch of morphological analysis may not be
appropriate for the case study at hand. 
In this paper we present Sarf, a case-based morphological 
analyzer for Arabic. 
Sarf allows a case-based controller to control and refine
the state of the analysis on the fly. 
It uses non-deterministic finite state Automata
to analyze a stream of text. 
%Each machine corresponds to a valid analysis. 
Sarf uses concatenative analysis based on recursive affixes 
to better preserve part of speech information.
We automated the analysis of three books of Islamic
narrations (hadith) using Sarf where we
abstracted each book into a vector of complex structures.
Our results show higher accuracy and better efficiency
compared to current analyzers.
\end{quote}
\end{abstract}

@@

%Automated analysis of Arabic data sets, including texts, 
%publications, records and digital media is essential
%with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
Current morphological analyzers~\cite{Sughaiyer:04}
use concatenative analysis when
considering the internal structure 
of an Arabic word and
composing it into several {\em morphemes}. 
A morpheme can be a {\em stem}, or an {\em affix}.
An affix can be a {\em prefix, suffix, } or an {\em infix}.
The analysis of one word may lead to several possible
morphological solutions.
\vocalize
For instance, the word \RL{'a.hmadH}~\footnote{In this document, we use the default ArabTeX transliteration style ZDMG.}
may have two valid morphological analyses. 
The letter \RL{'a} may be a prefix and the word means 
``I praise him'', or 
%The letter \RL{'a} may also be 
part of the stem \RL{'a.hmad} (a proper noun)
and the word means ``his Ahmad''.

\novocalize
The accuracy of the solutions suffer due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels. 
This greatly increases the ambiguity of Arabic text. 
Arabic letters can have up to %(maybe reduce this sentence)
four different forms
corresponding to their position in a word, i.e, beginning,
middle, end of word and separate forms. 
This allows the phrase \transfalse
\RL{il_A\nospace almdrsT} \transtrue
to be visually recognizable
as two separate words \RL{il_A} (to) and \RL{almdrsT} (the school) 
without the need of a space in between. 
The reason is the first word \RL{il_A} ends with
\RL{_A} a non-connecting letter. 
These words,
referred to as ``run-on'' words~\cite{Buckwalter:04},
occur often, and greatly increase the
difficulty of tokenization.

{\bf Sarf.~~~}
In this paper, we present Sarf~\footnote{Sarf (name modified for blind review) 
has been submitted with the paper and will be available online as an open source tool.},
 a {\em novel case-based efficient
morphological analyzer} that uses 
non-deterministic finite state Automata 
driven by an application-specific controller.
Each machine in Sarf takes one letter at a time as input
and represents a valid morphological analysis.
The application-specific controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
The controller we present in Section~\ref{sec:controller}
is an FSA. However, as we explain in Section~\ref{sec:ndfsa},
it can be a logical entity of any type.

%Each alive machine in Sarf takes as input one letter at a time 
%from a text stream. 
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
Up to our knowledge, this is the first morphological analyzer that 
handles the ``run-on'' words problem.

Sarf uses concatenative analysis with a novel 
{\em recursive affixes} refinement,
where an affix can be a sequence of other affixes which are concatenatively compatible.
%An atomic affix $a$ is a standalone affix
%that is not the concatenation of other affixes. 
%We define an affix $f: a | a f$ to be either
%an atomic affix $a$ or an atomic affix concatenated
%with an affix. 
Sarf introduces recursive affixes in order to
better retain part of speech information and enhance the 
efficiency of affix matching.

Our main focus when building Sarf, was not on improving the semantical rules and expanding the morphological alternatives, instead we worked on adding support for higher level control by the specific application utilizing the lower level morphological analyzer. In other words, application-specific controller, has full control over and directs the morphological analysis of Sarf. % (sentence should be reworded differently, but is needed to explain why we mention higher and lower accuracy hypothesis, which seems vague otherwise)

We hypothesize that many NLP case studies need the 
morphological analyzer to answer queries that do not need 
high accuracy at the low morphological analysis level.
The analysis at the higher levels 
(i.e. answering the application-specific query)
can often compensate for 
tolerable inaccuracy at lower morphological levels. 
For example, if a query concerns proper names and the 
analyzer is considering a 
prefix that connects only to a verb,
%the analyzer can answer without considering the 
%rest of the word.
we do not need to further analyze the rest of the word 
to provide a negative answer.

We find evidence to our hypothesis in~\cite{Maamouri:10}.
The addition of a new corpus to the Arabic Tree Bank 
(ATB)~\cite{Maamouri:04}
%required compliance with the DARPA GALE program 
%including several NLP tasks such as English translation 
%annotations and word-level alignment of Arabic and English. 
%The new data 
challenged the existing processing and 
annotation guidelines. 
This led to refining 
SAMA~\cite{Kulick:10} with 
alternative guidelines. 
This refinement was achieved only after the NLP 
task interacted with the morphological analyzer,
a flexibility our analyzer is designed to provide for the application-specific controller on the run. %(check if this sentence makes sense)
In addition, 
the work of Habash~\shortcite{Habash:06} reports that different 
types of morphological analyses behave differently on the same %case study. 
data set.
The analyzers may report different solutions or 
may present the solutions in different orders. 
Clearly, some of 
the presented solutions and their order of presentation may not be 
appropriate for the %case study.
application needs, which is an issue that Sarf tackles. % (maybe reworded)
Moreover, our results strongly support our claim that
an application-specific morphological analyzer that allows a application-specific 
controller to intervene at every decision to guide, use, order, and 
refine the morphological analysis is of high utility.
We validate our hypothesis using a case study from the Islamic 
literature, introduced later in Section~\ref{sec:islamic}.

The rest of this paper is structured as follows. In Section~\ref{sec:related}
we compare Sarf to related work. In Section~\ref{sec:sarf}
we discuss Sarf and illustrate it with a running example.
%We also explain how non-deterministic finite state Automata
%work. 
In Section~\ref{sec:islamic} we present the case
study. We present our results in Section~\ref{sec:results}
and conclude with future work in Section~\ref{sec:future}.

@@

\section{Related work }
\label{sec:related}

The survey in~\cite{Sughaiyer:04} compares
several morphological analyzers. 
Analyzers such as~\cite{Khoja:01,Darwish:02} 
target specific applications in the 
analyzer itself or use a specific set of POS tags
as their reference.
Sarf differs in that it is a general morphological 
analyzer that reports all possible solutions. 
It is application-specific in the sense that a controller 
prunes these solutions.

Current morphological analyzers such as 
Buckwalter~\shortcite{Buckwalter:02},
Beesley~\shortcite{Beesley:01},
SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} 
take as input white space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
%This approach has several problems. 
However, we note the following points that deserve consideration:
%(We need not to mention especially point 2 as a problem according to reviewer comments)
First, a white space delimited token may have 
more than one word.
%Second, the exhaustive enumeration may hurt performance and may
%not be necessary or appropriate
%in some case studies as noted in~\cite{Maamouri:10}. 
Second, in many applications, the exhaustive enumeration may not be needed or appropriate as noted in~\cite{Maamouri:10}; consequently, adding unnecessary running time.
Other morphological analyzers such as 
Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
% (reviewer wants to distinguish MADA as a tool for morphological disambiguation instead of analysis)
use machine learning and support vector machines (SVM) 
to enhance the accuracy of the morphological analysis at the expense 
of performance. % (we should stress that we mean performance in time and not accuracy although evident since some reviewers did not get)
Xerox rules can be compiled into specialized finite state
machine (FSM) based analyzers as described in~\cite{Beesley:03}.
However, the efficiency of the resulting analyzer depends on the
way the Xerox rules are written. 
This requires deep knowledge and insight from the user
in compilation techniques, context free grammars, 
and morphological analysis.
%context free grammar rules that can be automatically compiled
%into efficient non-deterministic finite state Automata (NDFSA).

Sarf builds upon the lexicon of Buckwalter\shortcite{Buckwalter:02}.
It differs from Buckwalter in that it defines a shorter list of affixes
and a longer list of affix compatibility rules to allow 
affixes to be defined recursively
so that we can have prefix-prefix an suffix-suffix 
concatenations.
This allows Sarf to better maintain and propagate 
the POS tags associated with affixes. 
%This information is highly 
%redundant with Buckwalter.% and that may lead to consistency issues. 
Sarf keeps all the possible analyses alive where each analysis
corresponds to a set of decisions and positions in the text. 
Buckwalter produces a set of solutions for each token 
and the several analysis threads of the text need to be 
generated as a product of the token solutions.

Similar to Sarf, Beesley~\shortcite{Beesley:01} uses
finite state Automata (FSA). 
It reports that the number of Automata generated by a compiler
for Xerox rules can not be controlled and also reports a difficulty 
to compose the FSAs into a single framework. 
Sarf however constructs an NDFSA from
the controller and the prefix, stem, and suffix deterministic 
transducers to solve the problem.
Sarf also differs in that it uses recursive affixes in the 
affix transducers, and in that the transducers consider
input from the controller to modify their behavior. 
%With Sarf it is easy to control the FSA with a case-based
%controller. 
Beesley requires recompilation of the Xerox rules. % (don't we do also for any change in the controller code?)

Similar to The MADA+TOKAN toolkit, Sarf thinks of
the several valid morphological analyses as a richness that 
should be exploited at a higher abstract level rather than
an inaccuracy that should be corrected. 
Sarf differs in that it provides an interface for the 
case study to interfere in judging the analyses much earlier and
at a finer granularity level. 
In Sarf, there is no need for a separate tokenizer such as
TOKAN since each solution keeps a stack of positions
that partition text into morphemes.
%The case-based controller can build the trace 
%when needed.

@@

The diagram in Figure~\ref{f:example}
illustrates the operation of $\Phi$, the Sarf FSA,
when parsing the
string \RL{wsyl`b-h-aa al-lA`bwn}.
%\RL{al-lA`bwn}\RL{wsyl`b-h-aa}.
\footnote{ 
\RL{n}%
\RL{w}
\RL{b}
\RL{`}
\RL{a} 
\RL{l}
\RL{l}
\RL{a} %
\RL{a}
\RL{h}
\RL{b}
\RL{`}
\RL{l}
\RL{y}
\RL{s}
\RL{w}
in separate form to ease following the example
}

@@

The diagram in Figure~\ref{f:composition} shows an 
abstraction of the non-deterministic construction of Sarf
from the prefix, stem and suffix transducers. 
The circle represents regular states, the square
represents accept states and the triangle represents
reject states. 
Note that the concrete prefix, stem and suffix transducers
are all deterministic and cycle free as discussed 
is Sections~\ref{sec:affixFSA} and~\ref{sec:stemFSA}.
The cycles are introduced in the abstract machine
for pedagogical purposes.

@@

\begin{table*}[bt]
\centering
\caption{Results of the hadith case study with Sarf.}
%\begin{tabular}{|p{1.5cm}||c|c||c|c||c|c|} \hline
\resizebox{1.9\columnwidth}{!}{
\begin{tabular}{lp{.2cm}ccp{.2cm}ccp{.2cm}cc} %\cline{2-10}
 & & \multicolumn{2}{c}{Al Kafi} & & \multicolumn{2}{c}{Al Istibsar} & &\multicolumn{2}{c}{Ibn Hanbal} \\ \cline{3-10}
 & & Simple & Refined & & Simple & Refined & & Simple  & Refined \\\bottomrule  %\hline \hline
Word count & &\multicolumn{2}{c}{98,943} & & \multicolumn{2}{c}{103,835} & & \multicolumn{2}{c}{20,354} \\ 
 Names      & & 20,473 & 12,060  & &20,646 &  14,613& & 4,762 & 3,013\\
Names/Narrator& &2.48 & 1.97 & & 2.17 &  1.84& & 1.49 & 1.25 \\
Narrators & &3,080 & 2,623 & & 6,164 &  5,767& & 2,082 & 1,755 \\ 
Narrators/Chain & & 5.28 & 4.84 & & 4.98&  4.76 & & 4.8 &4.05 \\
Chains & & 583 & 542 &  & 1,238& 1,211& & 433 & 433 \\ 
Ignored names & & 11,801 & 6,400 &  & 5,845 & 3,348 & & 1,386 & 642 \\ \hline
Segmentation accuracy & & 85\% & 96\%& & 96\%& 96\%& & 92\%& 92\%\\ 
Chain accuracy & & 99\% & 99\%&  & 98\%& 99\%& & 99\% & 97\% \\ 
%Narrator accuracy & & & & & &  & & & \\ 
Narrator name accuracy & & 93\% & 91\%& &  92\%& 90\% & & 90\% & 90\% \\ \hline
Name false positives & & 28\% & 7\%&  & 9\%& 4\% & & 34\% & 4\% \\ \hline
Sarf running time (secs.)& & 2.97 & 1.32 & & 2.93 & 1.31 & & .25 & .096\\ \hline \hline
 & & Buckwalter& ElixirFM & & Buckwalter & ElixirFM & & Buckwalter & ElixirFM \\\bottomrule
Names      & & 18,354 & 19,201 & & 17,934 &  18,202 & & 5,021 & 5,117 \\
Names/Narrator& & 2.6 & 2.1 & & 2.47 &  2.07 & & 1.83 & 1.62\\
Narrators & & 3,210 & 2,981 & & 6,425 &  6,208 & & 2,311 & 1,965 \\ 
Narrators/Chain & & 5.57 & 5.32 & & 5.46 &  5.32 & & 5.12 & 5.01 \\
Chains & & 627 & 601 &  & 1,318 & 1,250 & & 487 & 452 \\ 
Ignored names & & 9,642 & 9,915 &  & 6,118 & 6,874 & & 1,812 & 1,904 \\ \hline
Segmentation accuracy & & 76\%& 84\%& & 85\%& 89 \%& & 84\%& 85\%\\ 
Chain accuracy & & 78\% & 87\%&  & 87\%& 90\%& & 85\% & 86\% \\ 
%Narrator accuracy & & & & & &  & & & \\ 
Narrator name accuracy & & 76\% & 86\%& & 84\%& 87\% & & 84\% & 84\% \\ \hline
Name false positives & & 29\% & 22\%&  & 14\%& 13\% & & 36\% & 32\% \\ \hline
Running time (secs.)& & 6.65 & 2.78$\times60^2$& &4.55 & 2.3$\times60^2$ & & .66 & 29.2$\times$60 
\end{tabular}
}
\normalsize
\label{t:hadithresallresults}
\end{table*}

In this section we present results that compare Sarf to existing 
morphological analyzers such as Buckwalter and ElixirFM. 
We also evaluate our hypothesis of the utility of a case-based 
morphological analyzer.
We validate that the high-level analysis of the output 
of low-level morphological analysis can compensate for accuracy.
To achieve that we ran Sarf in two modes, simple and refined.
In the simple mode, we used Sarf as discussed in 
Section~\ref{sec:sarf} with the 
controller of Section~\ref{sec:controller}.
In the refined mode, we applied several low-level morphological 
refinements to Sarf such as disambiguation of 
female proper names so that the controller can ignore them. 
We also added a refinement that detected compound proper names 
such as \RL{`bd alr.hmn} 
as they are common in narrator names.
%Finally, we consider a possessive form as a reference to 
%a person and treat
%it as a narrator.

We ran our experiments using a dual core 2.66 Ghz 64-bit processor 
with 4GB of memory using a Linux operating system.

We report our results in Table~\ref{t:hadithresallresults} on the 
three hadith books.
In general, we observe that the refinements slightly 
improved performance and accuracy at
the lowest level of abstraction by reducing false positives.
However, the refinements had little to no effect on the high 
accuracy measures at higher levels of abstraction.
We believe that this is a strong supporting evidence to our 
hypothesis.

The row for ``word count'' is an indicator of the size of the data.
The %numbers in 
rows ``names'', ``narrators'', and ``chains'' denote the total
number of the corresponding entity per book.
The relative magnitudes of these rows
reflect the hierarchical structure where 
a chain contains
narrators, and a narrator contains names.
The row ``ignored names'' reports the number of names detected 
outside the chains and thus are irrelevant to
the case study.
In addition, the rows ``names/narrator'' and ``narrators/chain'' 
report the average
number of names per narrator and 
the average number of narrators per chain respectively.

The ``segmentation accuracy'' reports on the accuracy of 
segmentation of the book into separate hadiths.
The metric is computed manually over a representative sample 
of each book.
The same applies to all other accuracy measures.
The ``chain accuracy'' reflects the percentage of correct narrators 
covered by extracted chains.
The ``narrator name accuracy'' row reflects the percentage of 
correct proper names detected within
narrator structures.

The row ``name false positives'' reports the percentage of words 
recognized as names while they should not.
We notice that this low level metric improved dramatically 
with the refined analysis without substantially affecting the 
high abstract measures such as chain and segmentation
accuracy.

>>>>

\begin{document}
\maketitle
\begin{abstract}
Natural language processing (NLP) uses morphological 
analysis to abstract and annotate text.
Often times, 
%      and in particular in the context of the Arabic language, 
annotations resulting from 
a branch of morphological analysis may not be
appropriate for the application at hand. 
In this paper we present Sarf, 
%an application-specific morphological analyzer for Arabic. 
an Arabic morphological analyzer, which
%Sarf
allows an application-specific controller to control and refine
the state of the analysis on the fly. 
It uses non-deterministic finite state Automata % (NDFSA)
to analyze a stream of text. 
%Each machine corresponds to a valid analysis. 
Sarf uses concatenative analysis 
%based on
with  
recursive affixes 
to better preserve part of speech (POS) information.
It solves the issue of ``run-on'' words 
and utilizes diacritics for disambiguation.
%We automated the analysis of three books of Islamic
%narrations (hadith) using Sarf, where we
%abstracted each book into a vector of complex structures.
Our results show better efficiency
compared to current analyzers.
\end{quote}
\end{abstract}

@@

%Automated analysis of Arabic data sets, including texts, 
%publications, records and digital media is essential
%with the huge digital Arabic content available nowadays. 
Morphological analysis is key in current automated 
analysis techniques for Arabic text. 
%Current morphological analyzers~\cite{Sughaiyer:04}
%use concatenative analysis when
%considering the internal structure 
%of an Arabic word and
%composing it into several {\em morphemes}. 
%A morpheme can be a {\em stem}, or an {\em affix}.
%An affix can be a {\em prefix, suffix, } or an {\em infix}.
%The analysis of one word may lead to several possible
%morphological solutions.
%\vocalize
%For instance, the word \RL{'a.hmadH}~\footnote{In this document, we use the default ArabTeX transliteration style ZDMG.}
%may have two valid morphological analyses. 
%The letter \RL{'a} may be a prefix and the word means 
%``I praise him'', or 
%%The letter \RL{'a} may also be 
%part of the stem \RL{'a.hmad} (a proper noun)
%and the word means ``his Ahmad''.
%
%\novocalize
The accuracy of the solutions suffers due to inherent difficulties
of morphological analysis of the Arabic language. 
For example, it is common practice to write Arabic text
without short vowels (diacritics) 
which greatly increases its ambiguity.
%This greatly increases the ambiguity of %Arabic text.
%the text.
%Moreover, Arabic letters can have up to 
%four different forms
%corresponding to their position in a word, i.e, beginning,
%middle, end of word and separate forms. 
%This allows the phrase
%\noTrnoVocRL{il_A\nospace almdrsT}
%to be visually recognizable
Moreover, determining word boundaries is not always 
%a trivial problem.
trivial.
For example,
\noTrnoVocRL{il_A\nospace almdrsT}
can be visually recognizable
as two separate words \noVocRL{il_A} (to) and \noVocRL{almdrsT} (the school) 
without the need of a space in between. 
%The reason is the first word %\noVocRL{il_A} 
%ends with
%\noTrnoVocRL{_A} a non-connecting letter. 
The reason is that
\noTrnoVocRL{_A} is a non-connecting letter. 
%These words, referred to as 
Such 
``run-on'' words~\cite{Buckwalter:04},
%occur often, and greatly increase the
occur often, %greatly
increasing the
difficulty of tokenization.

%{\bf Sarf.~~~}
In this paper, we present Sarf~\footnote{Sarf (name modified for blind review) 
has been submitted with the paper and will be available 
%online
as an open source tool.}, 
a {\em novel application-specific efficient 
%(maybe 'enhanced' instead of 'efficient' to engulf both efficiency and other refinements)
morphological analyzer} that uses 
non-deterministic finite state Automata (NDFSA)
driven by an application-specific controller.
Each machine in Sarf takes one letter at a time as input
and represents a valid morphological analysis.
%The application-specific controller prunes false positives
The controller prunes false positives
early in the run by making a decision at every input letter
to control the machines.
%The controller we present in Section~\ref{sec:controller}
%is an FSA. However, as we explain in Section~\ref{sec:ndfsa},
%it can be a logical entity of any type.
The controller can be a Finite State Automaton (FSA) or any other %logical entity 
programming entity, % of any type,
%since Automatas may be a %very complex
%non-necessary solution for some simple applications.
%(maybe re-word this sentence)
depending on application needs.

%(check if position of this paragraph seems in correct place)
%Our focus when building Sarf, was not on improving the semantical rules and 
%expanding the morphological alternatives, instead we worked on adding support 
%for higher level control by the specific application utilizing the lower level 
%morphological analyzer. %In other words, the application-specific controller, has 
% full control over and directs the morphological analysis of Sarf. 
% (sentence should be reworded differently, but is needed to explain why we mention 
% higher and lower accuracy hypothesis, which seems vague otherwise)
%
%We have noted that many NLP case studies need the 
%morphological analyzer to answer queries that do not need 
%high accuracy at the morphological analysis level.
%The analysis at the higher levels 
%(i.e. answering the application-specific query)
%can often compensate for 
%tolerable inaccuracy at lower morphological levels. 
%For example, if a query concerns proper names and the 
%analyzer is considering a 
%prefix that connects only to a verb,
%%the analyzer can answer without considering the 
%%rest of the word.
%we do not need to further analyze the rest of the word 
%to provide a negative answer.
%
%We find evidence to our hypothesis in~\cite{Maamouri:10}.
%%The addition of a new corpus to the Arabic Tree Bank 
%%(ATB)~\cite{Maamouri:04}
%%%required compliance with the DARPA GALE program 
%%%including several NLP tasks such as English translation 
%%%annotations and word-level alignment of Arabic and English. 
%%%The new data 
%%challenged the existing processing and 
%%annotation guidelines. 
%%This led to refining 
%%SAMA~\cite{Kulick:10} with 
%%alternative guidelines. 
%%This refinement was achieved only after the NLP 
%%task interacted with the morphological analyzer,
%%a flexibility our analyzer is designed to provide for the application-specific controller on the run. %(check if this sentence makes sense)
%It reports that a refined version of SAMA~\cite{Kulick:10} was achieved when NLP task interacted with the morphological analyzer.
%%a flexibility our analyzer is designed to provide for the application-specific controller on the run.
%In addition, 
%the work of Habash~\shortcite{Habash:06} reports that different 
%types of morphological analyses behave differently on the same %case study. 
%data set.
%The analyzers may report different solutions or 
%may present the solutions in different orders. 
%Clearly, some of 
%the presented solutions and their order of presentation may not be 
%appropriate for the %case study.
%application needs, which is an issue that Sarf tackles. % (maybe reworded)
We have noted that in many NLP applications, when the morphological analyzer
benefits from the peculiarities and characteristics of the query at hand, 
the analysis %can be easily 
is greatly 
enhanced. 
Consequently, we strongly believe that by 
%being 
allowing
an application-specific %morphological 
%analyzer that allows %an application-specific 
%the
controller to intervene at every decision to guide, use, order, and 
refine the morphological analysis, Sarf is of high utility.

%Each alive machine in Sarf takes as input one letter at a time 
%from a text stream.
Apart from being application-specific, 
Sarf %has
%also  
introduced 
many enhancements. %to the field of Arabic morphological analysis. 
%the following enhancements:
%For example,
First,
Sarf does not assume that the word at hand is a token and
performs tokenization on the fly based on morphological correctness.
Up to our knowledge, this is the first %morphological 
analyzer that 
handles %the ``run-on'' words problem. 
``run-on'' words. 
%
%Moreover, since Sarf performs tokenization on the fly, Sarf is not just capable of solving
%the problem of ``run-on'' words, but also of handling Multiword Expressions (MWE). 
Moreover,
% since Sarf does not perform white-space delimited tokenization, 
Sarf 
%is capable of handling
handles 
Multiword Expressions (MWEs),
%MWEs are defined in~\cite{MWE} as
which are 
``idiosyncratic interpretations
that cross word boundaries (or spaces)''~\cite{MWE}. 
%By this term, we refer to frequent cases when a group of words form together a single
%item with a semantical meaning. In this paper, we focus on the example of compound names such as 
%\noVocRL{`bid alkarym}. 
%In current analyzers, this phrase is analyzed word by word and in best cases is 
%considered to contain two names, instead of being understood to be a single but compound name.

Second, 
Sarf uses concatenative analysis with a novel %(maybe remove novel bc authors did not feel it is much)
{\em recursive affixes} refinement,
%where an affix can be a sequence of other affixes which are concatenatively compatible.
where an affix can be a sequence of other concatenatively compatible affixes.
%An atomic affix $a$ is a standalone affix
%that is not the concatenation of other affixes. 
%We define an affix $f: a | a f$ to be either
%an atomic affix $a$ or an atomic affix concatenated
%with an affix. 
%Sarf introduces recursive affixes in order to
%This refinement is introduced to better retain
Hence, it better retains
%part of speech information
POS information  
and enhances the 
efficiency of affix matching.

Third, Sarf supports disambiguation using 
%diacritics when they are available.
any available diacritics.
%This feature, to the best of our knowledge is the first attempt to benefit from 
%To the best of our knowledge,
Up to our knowledge, this is the first attempt to benefit from
diacritics instead of ignoring them. 
Prior to Sarf, the trend was to preprocess 
the text to remove  
these %any such
`unwelcomed' letters. 
%Nonetheless, Sarf regards them as a richness to be utilized that should not be disregarded.

%In addition to all its enhanced analysis methods, Sarf still
%Sarf also
%outperformed current morphological analyzers in terms of running time,
%%These results will be presented 
%as shown in Section~\ref{sec:results}.
Sarf is also more efficent than current morphological analyzers, 
as shown in Section~\ref{sec:results}.

%%We make the following 
%We summarize our key contributions: 
%\begin{enumerate}
%\item {\bf application-specific analysis:}  We designed and implemented
%a application-specific morphological analyzer where a controller
%guides the analysis. 
%%\item {\bf Exhaustive analysis:} Since we have an application-specific 
%%controller that can decide on the fly whether an analysis is 
%%accepted or not, we can afford to keep all valid analyses modulo
%%those pruned by the case controller. 
%%We do that using non-deterministic FSAs. 
%\item {\bf Recursive affixes:}
%We refine the concatenative analysis of 
%Buckwalter~\shortcite{Buckwalter:02} and define affixes
%recursively. 
%This allows Sarf to perform better and 
%maintain compositional part of speech (POS) tags.
%\item {\bf Diacritic-based Disambiguation:}
%Diacritics are used to disambiguate morphological
%alternatives using any existant diacritics.
%\end{enumerate}

%The rest of this paper is structured as follows. In Section~\ref{sec:related}
%we compare Sarf to related work. In Section~\ref{sec:sarf}
%we discuss Sarf and illustrate it with a running example.
%%We also explain how non-deterministic finite state Automata
%%work. 
%In Section~\ref{sec:islamic} we present the case
%study. We present our results in Section~\ref{sec:results}
%and conclude with future work in Section~\ref{sec:future}.

@@

\subsection{Related work }
\label{sec:related}

The survey in~\cite{Sughaiyer:04} compares
several morphological analyzers. 
Analyzers such as~\cite{Khoja:01,Darwish:02} 
target specific applications 
%in the analyzer itself or use a specific set of POS tags as their reference
or limit themselves to a specific set of POS tags. 
Sarf differs in that it is a {\em general} morphological 
analyzer that reports all possible solutions. 
It is {\em application-specific} in the sense that a controller 
prunes these solutions.

Current morphological analyzers such as %(maybe disambiguators also added here according to reviewer comments)
Buckwalter~\shortcite{Buckwalter:02},
Beesley~\shortcite{Beesley:01},
SAMA~\cite{Kulick:10},
and ElixirFM~\cite{Otakar:07} 
take as input white-space delimited tokens~\cite{Kulick:10},
consider them as words,
and enumerate all possible solutions. 
This approach has several problems. 
%However, we note the following: %(I thought of this change not to include point 2 as a "problem" according to reviewer comments)
First, a white-space delimited token may have 
more than one word or just part of a MWE.
%Second, the exhaustive enumeration may hurt performance and may
%not be necessary or appropriate
%in some case studies as noted in~\cite{Maamouri:10}. 
Second, in many applications, exhaustive enumeration may not be needed or appropriate
~\cite{Maamouri:10}; consequently, adding unnecessary running time.

Other morphological analyzers such as 
Amira~\cite{Diab:07,Benajiba:07},
MAGEAD~\cite{Habash:05}, and MADA+TOKAN~\cite{Habash:09} 
% (reviewer wants to distinguish MADA as a tool for morphological disambiguation instead of analysis)
use machine learning and support vector machines (SVM) 
to enhance the accuracy of the morphological analysis at the expense 
of efficiency.
Xerox rules can be compiled into specialized 
%finite state machine (FSM) based analyzers as described in~\cite{Beesley:03}.
FSA-based analyzers~\cite{Beesley:03}.
However, the efficiency of the resulting analyzer depends on the
way the Xerox rules are written. 
This requires deep knowledge and insight from the user
in compilation techniques, context free grammars, 
and morphological analysis.
%context free grammar rules that can be automatically compiled
%into efficient non-deterministic finite state Automata (NDFSA). 
Sarf, however, allows users to write directly the code for the needed controller. %(maybe not clear)

Sarf builds upon the lexicon of Buckwalter. %\shortcite{Buckwalter:02}, %(already cited above)
%It differs from Buckwalter in that it defines a shorter list of affixes
It defines a shorter list of affixes
and a longer list of 
%affix 
compatibility rules to allow 
affixes to be defined recursively
so that we can have prefix-prefix and suffix-suffix 
concatenations.
%This allows Sarf to better maintain and propagate 
Hence, it better maintains and propagates 
the POS tags associated with affixes. 
%This information is highly 
%redundant with Buckwalter.% and that may lead to consistency issues. 
% (check if necessary to keep the 2 sentences below)
%Sarf keeps all the possible analyses alive where each analysis
%corresponds to a set of decisions and positions in the text. 
%Buckwalter produces a set of solutions for each token 
%and the several analysis threads of the text need to be 
%generated as a product of the token solutions.

Similar to Sarf, both Beesley %~\shortcite{Beesley:01} (already cited previously)
and Attia~\shortcite{Attia:05} use
%finite state Automata (FSA). 
FSA.
Beesley reports that the number of Automata generated by a compiler
for Xerox rules can not be controlled and also reports a difficulty 
to compose the FSAs into a single framework. 
Sarf however constructs an NDFSA from 
%the controller and 
the prefix, stem, and suffix deterministic 
transducers to solve the problem.
Sarf also differs in that 
%it uses recursive affixes in the affix transducers, and in that 
the transducers consider
input from the controller to modify their behavior. 
%With Sarf it is easy to control the FSA with a application-specific
%controller. 
%Beesley requires recompilation of the Xerox rules. 
% (I am not sure if it is ok to have the above removed)
Similar to Sarf, Attia handles MWEs, but with a more extensive coverage
~\cite{Attia:06,Attia:10}.
%Attia has researched extensively MWEs handling methods~\cite{Attia:06} and 
%automated extending their dictionary coverage~\cite{Attia:10}. 
% (I am not sure how clear this statement is)

Similar to The MADA+TOKAN toolkit, Sarf thinks of
the several valid morphological analyses as a richness that 
should be exploited at a higher abstract level rather than
an inaccuracy that should be corrected. 
Sarf differs in that it provides an interface for the 
application to interfere in judging the analyses much earlier and
at a finer granularity level. 
In Sarf, there is no need for a separate tokenizer such as
TOKAN since each solution keeps a stack of positions
that partition text into morphemes.
%The application-specific controller can build the trace 
%when needed.

@@

The diagram in Figure~\ref{f:example}
illustrates the operation of $\Phi$, the Sarf FSA,
when parsing the
string \noTrnoVocRL{wsyl`b-h-aa\nospace al--lA`bwn}.
%\RL{al-lA`bwn}\RL{wsyl`b-h-aa}.
%\footnote{ 
%%\noVocRL{n}%
%%\noVocRL{w}
%%\noVocRL{b}
%%\noVocRL{`}
%%\noVocRL{a} 
%%\noVocRL{l}
%%\noVocRL{l}
%%\noVocRL{a} %
%%\noVocRL{a}
%%\noVocRL{h}
%%\noVocRL{b}
%%\noVocRL{`}
%%\noVocRL{l}
%%\noVocRL{y}
%%\noVocRL{s}
%%\noVocRL{w}
%\noTrnoVocRL{n}%
%\noTrnoVocRL{w}
%\noTrnoVocRL{b}
%\noTrnoVocRL{`}
%\noTrnoVocRL{a} 
%\noTrnoVocRL{l}
%\noTrnoVocRL{l}
%\noTrnoVocRL{a} %
%\noTrnoVocRL{a}
%\noTrnoVocRL{h}
%\noTrnoVocRL{b}
%\noTrnoVocRL{`}
%\noTrnoVocRL{l}
%\noTrnoVocRL{y}
%\noTrnoVocRL{s}
%\noTrnoVocRL{w}
%in separate form% to ease following the example
%}

@@

The diagram in Figure~\ref{f:composition} shows an 
abstraction of the non-deterministic construction of Sarf
from the prefix, stem and suffix transducers. 
The circle represents regular states, the square
represents accept states and the triangle represents
reject states. 
Note that the concrete prefix, stem and suffix transducers
are all deterministic and cycle free as discussed 
in Sections~\ref{sec:affixFSA} and~\ref{sec:stemFSA}.
The cycles are introduced in the abstract machine
for pedagogical purposes.

%Note that Sarf does not rely on implementing its machines as 
%equivalent regular expressions due to  the following reasons ...
% (not necessary)

@@

\begin{table}[tb]
\centering
\caption{Comparsion of Running Time}
\resizebox{0.95\columnwidth}{!}{
%\begin{tabular}{lp{.01cm}cp{.01cm}cp{.01cm}c}
\begin{tabular}{lccc}
\hline
 & Al Kafi & Al Istibsar & Ibn Hanbal \\ \cline{2-4} 
Sarf & 2.7002645 & 2.6884225 & 0.2356561 \\
Buckwalter & 3.315950 & 5.406016 & 0.5472224 \\
ElixirFM & 166.8$\times60$ & 138$\times60$ & 29.2$\times$60 \\
\hline
\end{tabular}
}
\normalsize
\label{t:efficeincyresults}
\end{table}

In this section we compare Sarf to existing 
morphological analyzers such as Buckwalter and ElixirFM in terms of running time.
We ran the different analyzers on
three books of Islamic narrations \RL{.hady_t} ~\cite{IbnHanbal,AlTousi,AlKulayni}
We report our results in Table~\ref{t:efficeincyresults} on the 
three hadith books.

Sarf outperforms both Buckwalter and ElixirFM in terms of 
running time. 
Buckwalter performed much better than ElixirFM in 
terms of running time, 
We think the big difference is due to the sophistication of 
the formal functional morphological analysis of ElixirFM 
and its reliance on additional annotations extracted from 
both ATB~\cite{Maamouri:04} and the Prague Arabic Dependency Treebank~\cite{Prague04}.
